---
title: "mandass2"
author: "Malvin"
date: "2022-09-28"
output:
  html_document: default
  pdf_document: default
  word_document: default
editor_options:
  markdown:
    wrap: 72
---

**Problem 1**\
*a)*\
Because $|x|$ is $-x$ fox $x<0$ and $x$ for $x>0$ we det the cumulative
distribution for $f(x)$ is:

$$ F(x)=\int_{x=-\infty}^x\frac12 e^{x}dx=\frac12e^x, \mbox{for} \ x\in \langle-\infty,0] $$
$$ F(x)=\int_{x=-\infty}^0\frac12 e^{x}dx +\int_0^x\frac12e^{-x}dx=1-\frac12e^{-x},\mbox{for} \ x\in[0,\infty \rangle. $$

To get the cumulative inverse, we have to set:

$$ F(x)=\frac12e^x=u, when \ x<0. $$ This is in
$u\in \langle 0,\frac12\rangle$. We then get: $$ e^x=2u$$
$$ F^{-1}(u)=ln(2u),u\in\langle0,\frac12\rangle $$\
\
And for positive x $$ F(x)=1-\frac12e^{-x}=u, \mbox{when} \ x>0 $$ This
is in $u\in\langle\frac12,1\rangle$. We then get: $$ 2-2u=e^{-x} $$
$$ F^{-1}(u)=-ln(2-2u), \ u\in\langle\frac12,1\rangle $$.

This means that the cumulative inverse can be written as:\
$$ F^{-1}(u) = \left\{ \begin{array}{}
         ln(2u), & \mbox{if $0<u<\frac12 $}\\
        -ln(2-2u), & \mbox{if $\frac12<u<1$}.\end{array} \right. $$\
\
*b)*\

```{r}
rm(list=ls())
start1 <- Sys.time() #Noting the time when we start the function
simfun <- function(n){
  u <- runif(n)
  dist <- c(log(2*(u[u<0.5])),-log(2-2*(u[u>0.5])))#Creating a vector with the different inverse functions for F. With u we get the distribution we want. All u-values are used, and only in the right function.
}
hist(simfun(100000), main="Histogram over simulation",breaks=250,probability=TRUE, density=10,xlim=c(-5,5), xlab="X-values")#Doing the simulation with 1000 values and plotting in a histogram. 
stop1 <- Sys.time() #Noting the time when we stop the function
```

\
\
*c)*\

```{r}
start2 <- Sys.time()
simalt <- function(n){
  y <- sample(c(-1,1),size=n, replace=TRUE)#Makes values of -1 and 1, which has 0.5 probability of each
  g <- rexp(n,1) #Exponentially distributed with rate 1
  dist2 <- y*g #Using information in problem to make the dist
}
hist(simalt(100000), main="Histogram over simulation",probability=TRUE,breaks=250, density=10,xlim=c(-5,5), xlab="X-values")
stop2 <- Sys.time()
time1 <- stop1-start1 #Substracting to find the difference in time
time2 <- stop2-start2

print(paste0("The first function uses ", time1, " and the second one uses ", time2))
```

The time it uses to run is very similar, and I can't get an answer on
which is faster. Therefore I would use the second one, because it is
simpler to use if we first have to compute the inverse of the cumulative
distribution on the first one.\
\
*d)*\
If we try with this normal distribution, we first have that the function
$g(x)$ is then (Normal distributed with $\mu=0\ and\ var(X)=\sigma^2$:
$$ g(x)=\frac1{\sigma \sqrt{2\pi}}e^{-\frac{(x-\mu)^2}{2\sigma^2}}=\frac{e^{-\frac{x^2}{2\sigma^2}}}{\sigma\sqrt{2\pi}} $$\
We then see if we divide $f(x)$ on $g(x)$, we get:
$$ \frac{f(x)}{g(x)}=\frac{\frac12e^{-|x|}}{\frac{e^{-\frac{x^2}{2\sigma^2}}}{\sigma\sqrt{2\pi}}}=\sigma\sqrt{2\pi}\frac12e^{-|x|-(-\frac{x^2}{2\sigma^2})}=\sigma\sqrt{2\pi}\frac12e^{-|x|+\frac{x^2}{2\sigma^2}} $$\
Which we see go to infinity when $x=\pm\infty$, because the power of e
becomes increasingly larger with x after a certain point. Therefore we
can't find a $c$ that can fulfill $\frac{f(x)}{g(x)}\le c$\
\
\
**Problem 2**\
*a)*\
The values we use for $\lambda_{max}$ is when it is the largest. Since $\lambda_1$ is a function of cos and a constant, we know that it is the largest when $t=0$, which gives:
\[ \lambda_{1,max}=1+\frac12cos0=\frac32. \]
\
For the other function we get that the max lambdavalue depends on the stoptime. We then get:

\[ \lambda_{2,max}=\frac{1+10}6=\frac{11}6,\ for\ t>10 \]

\[ \lambda_{2,max}=\frac{1+t}6,\ for\ t[0,10] \].


```{r}
NHPPfun1 <- function(stoptime){
  lambdamax <- 3/2 
  expnumber <- lambdamax*stoptime  # Expected events for lambdamax in time given
  n <- 3*expnumber  # To be sure we simulate more than expected values
  tbetween <- rexp(n,lambdamax) # Times between arrivals
  if(tbetween[1]>stoptime){ # If stoptime is less than the first time for the even(no events in time interval), we give an empty vector
    NHPP <- c()
  }
  else{
  cumtimes <- cumsum(tbetween)   # Arrival times
  cumtimes <- cumtimes[cumtimes<=stoptime] # Only values less than stoptime
  nevents <- length(cumtimes) # Count the number of events
  u <- runif(nevents) # Uniformly distributed values
  NHPP <- c()
  
  for(i in 1:nevents){ # For all the values we try the function and put it in the new vector NHPP
    if(u[i]<=(1+1/2*cos(cumtimes[i]*pi/5))/lambdamax){
       NHPP <- c(NHPP,cumtimes[i])
      }
    }
  }
  return(NHPP)
}

NHPPfun2 <- function(stoptime){
  if(stoptime>10){ # If t is larger than 10, we get lambdavalue for 10, otherwise we have to use the t to find it.
    lambdamax <- 11/6
  }
  else{
    lambdamax <- (1+stoptime)/6
  }
  expnumber <- lambdamax*stoptime  # Same approach as for NHPPfun1, until for loop.
  n <- 3*expnumber  
  tbetween <- rexp(n,lambdamax) 
  if(tbetween[1]>stoptime){
    NHPP <- c()
  }
  else{
  cumtimes <- cumsum(tbetween)   
  cumtimes <- cumtimes[cumtimes<=stoptime] 
  nevents <- length(cumtimes)
  u <- runif(nevents)
  NHPP <- c()
  for(i in 1:nevents){
    if(u[i]<=10){ # The function is different before and after t=10. Therefore we take the function with t before
      if(u[i]<=((1+cumtimes[i]))/(6*lambdamax)){
         NHPP <- c(NHPP,cumtimes[i])
      }
    }
    else
         NHPP <- c(NHPP,cumtimes[i])
    
      
    }
  }
  return(NHPP)
}

k <- 0
l <- 0
for(i in 1:100){ # Running the functions 100 times and find the mean of the length for both of them
k <- k+length(NHPPfun1(10))
l <- l+length(NHPPfun2(10))
}
print(k/100)
print(l/100)
```

We then get that the expected values are about 10, which in this case is
what it should be.\
\
*b)*\

```{r}
k <- 0
l <- 0
nsim <- 1000
for(i in 1:nsim){ # Running the functions 1000 times and find the mean of the length for both of them
k <- k+length(NHPPfun1(5))
l <- l+length(NHPPfun2(5))
}
print(paste0("Estimate of E(N_1)= ", k/nsim))
print(paste0("Estimate of E(N_2)= ", l/nsim))

T3 <- matrix(nrow=2,ncol=1000)
for(i in 1:1000){
T3[1,i] <- NHPPfun1(30)[3]
T3[2,i] <- NHPPfun2(30)[3]
}

plot
hist(T3[1,1:100], col="#1B98E080",density=25, breaks=20,xlim=c(0,8), probability=TRUE, main="Histogram of 3")
hist(T3[2,1:100],add=TRUE, col="#1B98E00D", breaks=20, probability=TRUE)

```
\
\
*d)*\
```{r}
NHPPfun3 <- function(b){
  t <- 0
  x <- c()
  while(t<b){
    t <- sqrt(t^2+20*rexp(1,1)+20*t+100)-10
    if(t<b){
      x <- c(x,t)
    }
  }
  return(x)
}
y <- c()
for(i in 1:1000){
   y <- c(y,length(NHPPfun3(20)))
}
z <- rpois(1000,40)
hist(y,probability=T, breaks=25, col=rgb(1,0.97,0.8,0.5))
hist(z,probability=T, breaks=25, col=rgb(0.5,0.5,0.9,0.5),add=T)

```

\
\
\
**Problem 3**\
*a)*

```{r}
markov <- function(mu, fi, sigma,t){# Function which takes in the four different variables.
  k <- 1 
  x <- vector(length=t)
  x[1] <- rnorm(1,mu,sigma^2) # First X is normal distributed N(mu,sigma^2)
  while(k<t){
    x[k+1] <- mu+fi*(x[k]-mu)+(rnorm(1,0,(1-fi^2)*sigma^2)) # Putting in the function for X
    k <- k+1
  }
  return(x)
}
```

\
\
*b)*\

```{r}
nsim <- 1000
M1<-M2 <- matrix(nrow=3,ncol=nsim)# Making matrices with 3 rows and 500 columns 
for(i in 1:3){ 
  for(j in 1:nsim){
  M1[i,j] <- sum(markov(0,-0.8+0.8*(i-1),1,1000))/1000 # Get three columns with fi-values given in problem
  M2[i,j] <- sum((markov(0,-0.8+0.8*(i-1),1,1000))^2)/1000
  }
}
print(paste0("The mean of M1 with fi-values -0.8,0, and 0.8 respectively is ", mean(M1[1,])," ",mean(M1[2,])," and ",mean(M1[3,])))
print(paste0("The variance of M1 with fi-values -0.8,0, and 0.8 respectively is ", var(M1[1,])," ",var(M1[2,])," and ",var(M1[3,])))

print(paste0("The mean of M2 with fi-values -0.8,0, and 0.8 respectively is ", mean(M2[1,])," ",mean(M2[2,])," and ",mean(M2[3,])))
print(paste0("The variance of M2 with fi-values -0.8,0, and 0.8 respectively is ", var(M2[1,])," ",var(M2[2,])," and ",var(M2[3,])))
```



